Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2017-12-21 paper-ieee-trans.tex

\documentclass[journal]{IEEEtran}
\usepackage{graphicx}
\usepackage{array}

@@

@@

As DevOps suggests, frequent releases positively affects the reliability of the
software. The software updates of INDIGO--DataCloud products taking place since
the second major release are passing through a Continuous Delivery (CD)
pipeline that adds the packaging of the software right after the execution of
the quality checks (as part of the CI).

The steps defined within the CD
pipeline differ whether the software is to be distributed in the form of Docker
images or via operating system’s packages. In the latter case an extra
validation step is added, consisting in the product’s deployment using a
Configuration Management (CM) solution. The installation uses the packages
built in the previous step and uploaded to a testing
repository. In the case of Docker images, the CM is used to build the image
itself, installing and configuring the product before uploading it to the
DockerHub repository \cite{indigo-dockerhub}. Figure \ref{fig:fig_CD} shows
the workflow followed by the products distributed as Docker images.

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{images/devops.png}
\caption{Continuous Delivery workflow for Docker images.}
\label{fig:fig_CD}
\end{figure*}

@@

The requirements are applied at early stages in the software development
process, integrated in the Jenkins CI service \cite{indigo-jenkins}, so that any
bug or design issue is likely to be detected and corrected promptly. The source
code is publicly available in GitHub repositories under an organization called
{\sl indigo-dc} \cite{indigo-github} to increase the visibility of the product
catalogue, promoting the external contributions and software adoption. Furthermore,
the source code is compliant with community de--facto style standards,
selected from the variety of programming languages being used as seen in
Fig. \ref{fig:fig_codestyle}.

\begin{figure}[!t]
\centering
\includegraphics[width=0.5\textwidth]{images/codestyle.png}
\caption{Code style standards followed by INDIGO--DataCloud's software products.}
\label{fig:fig_codestyle}
\end{figure}

Unit testing coverage has increased between the first ({\sl INDIGO-1}) and the
second ({\sl INDIGO-2}) INDIGO--DataCloud releases, reaching an average value of
66.69\% on the {\sl INDIGO-2} release, very close to the recommended 70\% threshold
defined by the project’s SQA base criteria \cite{indigo-d31}. A comparison graph
between the code coverage for the products involved both in {\sl INDIGO-1} and
{\sl INDIGO-2} releases is shown in Fig. \ref{fig:fig_unittest}. As it can be seen,
a very small set of products with low coverage values highly impacted the overall
result, although a significant part of these products are upstream contributions
or products with codebases not fully under the control of the project. Nevertheless,
60\% of the products were over this threshold while 80\% of them have values above
the 50\% coverage.

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{images/unittest.png}
\caption{Unit testing coverage for the two major INDIGO--DataCloud releases.}
\label{fig:fig_unittest}
\end{figure*}

Fig. \ref{fig:fig_confman} shows the number of products that offer an automated means
for deployment. It shows an increase after the {\sl INDIGO-1} release with specially
in the previous weeks of each of the two major releases.

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth, height=50mm]{images/confman.png}
\caption{Trend line showing the adoption of Configuration Management tools throughout the project lifetime.}
\label{fig:fig_confman}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{images/disvis.png}
\caption{DevOps pipeline to distribute Docker images for Disvis application.}
\label{fig:fig_disvis}
\end{figure*}

@@

Two pilot infrastructures are at the disposal of developers and scientific
communities involved in the project. The aim of these testbeds is to test the
level of integration between the components involved in the INDIGO--DataCloud
solution and use cases validation, by deploying and executing the applications
with the last stable version of the software. A map of the Pilot Preview
infrastructure is depicted in Fig. \ref{fig:fig_pilotpreview}, it shows the
resources providers and the components or services they deployed and supported.

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{images/pilotpreview.png}
\caption{Resource centers supporting the Pilot Preview testbed and corresponding
set of deployed INDIGO-DataCloud components or services.}
\label{fig:fig_pilotpreview}
\end{figure*}

>>>>

\documentclass[journal]{IEEEtran}
\usepackage{graphicx}
\usepackage{array}
\usepackage{subfigure}

@@

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{images/devops.png}
\caption{Continuous Delivery workflow for Docker images.}
\label{fig:fig_CD}
\end{figure*}

@@

As DevOps suggests, frequent releases positively affects the reliability of the
software. The software updates of INDIGO--DataCloud products taking place since
the second major release are passing through a Continuous Delivery (CD)
pipeline that adds the packaging of the software right after the execution of
the quality checks (as part of the CI). 
The steps defined within the CD
pipeline differ whether the software is to be distributed via Docker
images (see Figure \ref{fig:fig_CD}) or via operating system’s packages. In the latter case an extra
validation step is added, consisting in the product’s deployment using a
Configuration Management (CM) solution. The installation uses the packages
built in the previous step and uploaded to a testing
repository. In the case of Docker images, the CM is used to build the image
itself, installing and configuring the product before uploading it to the
DockerHub repository \cite{indigo-dockerhub}.

@@

The requirements are applied at early stages in the software development
process, integrated in the Jenkins CI service \cite{indigo-jenkins}, so that any
bug or design issue is likely to be detected and corrected promptly. The source
code is publicly available in GitHub repositories under an organization called
{\sl indigo-dc} \cite{indigo-github} to increase the visibility of the product
catalogue, promoting the external contributions and software adoption. Furthermore,
the source code is compliant with community de--facto style standards,
selected from the variety of programming languages being used as seen in
Figure \ref{fig:fig_codestyle}.

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{images/codestyle.png}
\caption{Code style standards followed by INDIGO--DataCloud's software products.}
\label{fig:fig_codestyle}
\end{figure}

Unit testing coverage has increased between the first ({\sl INDIGO-1}) and the
second ({\sl INDIGO-2}) INDIGO--DataCloud releases, reaching an average value of
66.69\% on the {\sl INDIGO-2} release, very close to the recommended 70\% threshold
defined by the project’s SQA base criteria \cite{indigo-d31}. A comparison graph
between the code coverage for the products involved both in {\sl INDIGO-1} and
{\sl INDIGO-2} releases is shown in Figure \ref{fig:fig_unittest}. As it can be seen,
a very small set of products with low coverage values highly impacted the overall
result, although a significant part of these products are upstream contributions
or products with codebases not fully under the control of the project. Nevertheless,
60\% of the products were over this threshold while 80\% of them have values above
the 50\% coverage.

@@

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{images/unittest.png}
\caption{Unit testing coverage for the two major INDIGO--DataCloud releases.}
\label{fig:fig_unittest}
\end{figure*}

Figure \ref{fig:fig_confman} shows the number of products that offer an automated means
for deployment. It shows an increase after the {\sl INDIGO-1} release with specially
in the previous weeks of each of the two major releases.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth, height=50mm]{images/confman.png}
\caption{Trend line showing the adoption of Configuration Management tools throughout the project lifetime.}
\label{fig:fig_confman}
\end{figure}

\begin{figure*}	
	\centering
	\begin{subfigure}
		\centering
\includegraphics[width=\textwidth]{images/disvis.png}
\caption{DevOps pipeline to distribute Docker images for Disvis application.}
\label{fig:fig_disvis}		
	\end{subfigure}
	\quad
	\begin{subfigure}
		\centering
\includegraphics[width=\textwidth]{images/pilotpreview.png}
\caption{Resource centers supporting the Pilot Preview testbed and corresponding
set of deployed INDIGO-DataCloud components or services.}
\label{fig:fig_pilotpreview}
	\end{subfigure}
\end{figure*}

@@

@@

Two pilot infrastructures are at the disposal of developers and scientific
communities involved in the project. The aim of these testbeds is to test the
level of integration between the components involved in the INDIGO--DataCloud
solution and use cases validation, by deploying and executing the applications
with the last stable version of the software. A map of the Pilot Preview
infrastructure is depicted in Figure \ref{fig:fig_pilotpreview}, it shows the
resources providers and the components or services they deployed and supported.
